---
title: "Functions"
output: html_document
date: "2024-03-12"
---

---
title: "Functions"
output: html_document
date: "2024-01-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(knitr)
library(gridExtra)
```

## Making the deck
```{r}
makeOD <- function(){
suits <- c("Spades", "Clubs", "Diamonds", "Hearts")
Suit <- unlist(map(suits, rep, 13))
ranks <- c("A","K", "Q", "J", "10", "9", "8",
           "7", "6", "5", "4", "3", "2")
Rank <- rep(ranks, 4)
values <-c(11, 10, 10, 10, 10:2)
Value <- rep(values, 4)
ordered_deck <- data.frame(Rank, Suit, Value)
return(ordered_deck)
}
```

## Deck Functions
```{r}
shuffle <- function(deck) {
  sample_n(deck, nrow(deck), replace = F)
}

##Supposedly, more decks means an increase in house advantage
cloneDeck <- function(deck, n){
  deck1 <- deck
  numClones <- n-1
  for(i in 1:numClones){
    deck <- rbind(deck, deck1) 
  }
  return(deck)
}

penetration <- function(shufDeck, proportion){
  nrow <- nrow(shufDeck)
  lastp <- nrow-ceiling(nrow*proportion)
  pen <- sample((nrow/2):lastp, 1)
  shoeEnd <- tail(shufDeck, nrow-pen)
  shufDeck <- head(shufDeck, pen)
  deck_p <- list(shufDeck, shoeEnd)
  return(deck_p) 
}
```

##Game functions
```{r}
deal <- function(playerHand, dealerHand, shufDeck, shoe) {
  #Blackjack is dealt a specific way (one to player, one to dealer, 
  #one to player, one to dealer), which may not be necessary to capture in 
  #the simulation but it makes it feel more real
  newValues <- draw(playerHand, shufDeck, shoe) 
  playerHand <- newValues[[1]]
  shufDeck <- newValues[[2]]
  shoe <- newValues[[3]]
  
  newValues <- draw(dealerHand, shufDeck, shoe)
  dealerHand <- newValues[[1]]
  shufDeck <- newValues[[2]]
  shoe <- newValues[[3]]
  
  newValues <- draw(playerHand, shufDeck, shoe) 
  playerHand <- newValues[[1]]
  shufDeck <- newValues[[2]]
  shoe <- newValues[[3]]
  
  newValues <- draw(dealerHand, shufDeck, shoe)
  dealerHand <- newValues[[1]]
  shufDeck <- newValues[[2]]
  shoe <- newValues[[3]]
  
  ##Slight chance to be dealt double aces, so make sure they don't bust
  dealerHand <- aceCheck(dealerHand)
  
  dealtHands <- list(playerHand, dealerHand, shufDeck, shoe)
  return(dealtHands)
}

draw <- function(hand, deck, shoe) {
  if(nrow(deck)==0){
    newCard <- shoe[1,]
    hand <- rbind(hand, newCard)
    shoe <- shoe[-1,]
  }else{
  newCard <- deck[1,]
  hand <- rbind(hand, newCard)
  deck <- deck[-1,]
  } 
  newValues <- list(hand, deck, shoe)
  return(newValues)
}

##If a hand has an ace (valued at 11) and a total value over 21,
##then the ace should be valued at 1. 
aceCheck <- function(hand){
  handTotal <- sum(hand$Value)
  if(nrow(hand)<=1){
    return(hand)
  }else if(hand[1,]$Value=="11" && hand[2,]$Value=="11"){
    hand[2,3] <- 1
  }else if(handTotal >21){
  if((any(hand==11))==TRUE){
    rowAce <- match(11, hand$Value)
    hand[rowAce,]$Value <- 1
    }
  }
  return(hand)
}

dealerTurn <- function(hand, deck, shoe){
  handTotal <- sum(hand$Value)
  while(handTotal<17){
    newValues <- draw(hand, deck, shoe) 
    hand <- newValues[[1]]
    deck <- newValues[[2]]
    shoe <- newValues[[3]]
    hand <- aceCheck(hand)
    handTotal <- sum(hand$Value)
  }
  ##A soft hand is any hand that contains an ace that is counted as 11.
#It is common practice for casinos to allow dealers to hit on a soft 17
#Apparently, this increases the house advantage
#This doesn't have to be a while command, since no matter what card is drawn,
#The total will always be 18 or over, and thus a stay
  if(handTotal==17 && any(hand == 11) == TRUE){
    newValues <- draw(hand, deck, shoe) 
    hand <- newValues[[1]]
    deck <- newValues[[2]]
    shoe <- newValues[[3]]
    hand <- aceCheck(hand)
    while(handTotal<17){
    newValues <- draw(hand, deck, shoe) 
    hand <- newValues[[1]]
    deck <- newValues[[2]]
    shoe <- newValues[[3]]
    hand <- aceCheck(hand)
    handTotal <- sum(hand$Value)
  }
  }
  endTurnValues <- list(hand, deck, shoe)
  return(endTurnValues)
}

outcomeHands <- function(pHandT, dHandT){
  if(dHandT>21){
    outcome <- "Win"
    reason <- "Dealer Bust"
    result <- list(outcome, reason)
    return(result)
  } else if (pHandT > dHandT){
    outcome <- "Win"
    reason <- "Better hand"
    result <- list(outcome, reason)
    return(result)
  } else if (pHandT < dHandT){
    outcome <- "Loss"
    reason <- "Worse hand"
    result <- list(outcome, reason)
    return(result)
    } else{
    outcome <- "Push"
    reason <- "Tied hands"
    result <- list(outcome, reason)
    return(result)
    }
}
```

##Mimic the dealer game
```{r}
simMimicDealer <- function(numSims, numDecks){
mimicStats <-data.frame()
##Make deck, shuffle and determine shoe
ordered_deck <- makeOD()
newDeck <- cloneDeck(ordered_deck, numDecks)
shufDeck <- shuffle(newDeck)
deck_p <- penetration(shufDeck, 0.20)
shufDeck <- deck_p[[1]]
shoe <- deck_p[[2]]
for(i in 1:numSims){
    endRound <- mimicDealer(shufDeck, shoe)
    finalRow <- compileOutcome(endRound)
    mimicStats <- rbind(mimicStats, finalRow)
    shufDeck <-endRound[[5]]
    shoe <- endRound[[6]]
    ##If the deck ever reaches the shoe, create a new shuffled deck after the round has ended.
  if(nrow(shufDeck)==0){
    shufDeck <- shuffle(newDeck)
    deck_p <- penetration(shufDeck, 0.20)
    shufDeck <- deck_p[[1]]
    shoe <- deck_p[[2]]
    ##Otherwise update deck based on round
  }
}
return(mimicStats)
}

mimicDealer <- function(shufDeck, shoe){
  playerHand <- data.frame()
  dealerHand <- data.frame()
  dealtHands <- deal(playerHand, dealerHand, shufDeck, shoe)
  playerHand <- dealtHands[[1]]
  dealerHand <- dealtHands[[2]]
  shufDeck <- dealtHands[[3]]
  shoe <- dealtHands[[4]]
  pHandT <- sum(playerHand$Value)
  dHandT <- sum(dealerHand$Value)
  ## Blackjack check
  #if the player or dealer has an opening hand that contains both an Ace and a
  #card valued at 10 (Face cards and 10's), they automatically win. 
  #If both the player and dealer are dealt Blackjack, it's a push. 
  #Whether or not the game ends if the dealer is dealt a blackjack depends 
  #on the rules of the casino (since the dealer may peek at the face down card,
  #then turn it face up if it completes a blackjack)
  if(pHandT == 21){
    if(dHandT == 21){
      outcome <- "Push"
      reason <- "Double Blackjacks"
    }else{
      outcome <- "Win"
      reason <- "Blackjack"
    }
  }else if(dHandT == 21){
    outcome <- "Loss"
    reason <- "Dealer Blackjack"
  } else{
  ## if no BJ then play game normally
  ##Player turn (mimicking dealer's turn)
  endTurnValues <- dealerTurn(playerHand, shufDeck, shoe)
  playerHand <- endTurnValues[[1]]
  shufDeck <- endTurnValues[[2]]
  shoe <- endTurnValues[[3]]
  pHandT <- sum(playerHand$Value)
  ##if player busted, dealer wins without taking turn, 
  if(pHandT>21){
    outcome <- "Loss"
    reason <- "Bust"
  } else{
  #otherwise, dealer takes a turn
  endTurnValues <- dealerTurn(dealerHand, shufDeck, shoe)
  dealerHand <- endTurnValues[[1]]
  shufDeck <- endTurnValues[[2]]
  shoe <- endTurnValues[[3]]
  dHandT <- sum(dealerHand$Value)
  outcomes <- outcomeHands(pHandT, dHandT)
  outcome <- outcomes[[1]]
  reason <- outcomes[[2]]
  }
  }
  endRound <- list(playerHand, dealerHand, outcome, reason, shufDeck, shoe)
  return(endRound)
}
```

##Fixed Stay
```{r}
fixedStay <- function(pCard1, pCard2, dCard1, numDecks, deckPenetration){
OD <- makeOD()
OD_Stay <- shuffle(cloneDeck(OD, numDecks))
deck_p <- penetration(OD_Stay, deckPenetration)
OD_Stay <- deck_p[[1]]
shoe <- deck_p[[2]]
playerHand <- data.frame()
dHand <- data.frame()

C1Row <- match(pCard1, OD_Stay$Value)
C1 <- OD_Stay[C1Row,]
playerHand <- rbind(playerHand, C1)
OD_Stay <- OD_Stay[-C1Row,]

C2Row <- match(pCard2, OD_Stay$Value)
C2 <- OD_Stay[C2Row,]
playerHand <- rbind(playerHand, C2)
OD_Stay <- OD_Stay[-C2Row,]

C3Row <- match(dCard1, OD_Stay$Value)
C3 <- OD_Stay[C3Row,]
dHand <- rbind(dHand, C3)
OD_Stay <- OD_Stay[-C3Row,]

pHandT <- sum(playerHand$Value)
fixedStandResults <- data.frame()
for(i in 1:1000){
  SD1 <- shuffle(OD_Stay)
  dealerHand <- dHand
  endTurnValues <- dealerTurn(dealerHand, SD1, shoe)
  dealerHand <- endTurnValues[[1]]
  dHandT <- sum(dealerHand$Value)
  outcomes <- outcomeHands(pHandT, dHandT)
  outcome <- outcomes[[1]]
  reason <- outcomes[[2]]
  outcomeList <- list(playerHand, dealerHand, outcome, reason)
  finalRow <- compileOutcome(outcomeList)
  fixedStandResults <- rbind(fixedStandResults, finalRow)
}
Results <- fixedStandResults %>% group_by(outcome) %>% summarise(n=n(),
                                                           prob_of_occuring = n/1000)
WinResults <- fixedStandResults %>% group_by(outcome) %>% summarise(n=n(),
                                                           prob_of_occuring = n/1000) %>% filter(outcome=="Win")
LossResults <-fixedStandResults %>% group_by(outcome) %>% summarise(n=n(),
                                                           prob_of_occuring = n/1000) %>% filter(outcome=="Loss")
WinStay <- round(WinResults[1,3], 4)
LoseStay <- round(LossResults[1,3], 4)
expectation <- WinStay-LoseStay
probList <- list(expectation, WinStay, LoseStay)
listRes <- list(expectation, probList, fixedStandResults)
return(listRes)
}
```

##Fixed Hit
```{r}
fixedHit <- function(pCard1, pCard2, dCard1, numDecks, deckPenetration){
OD <- makeOD()
OD_Hit <- shuffle(cloneDeck(OD, numDecks))
deck_p <- penetration(OD_Hit, deckPenetration)
OD_Hit <- deck_p[[1]]
shoe <- deck_p[[2]]
pHand <- data.frame()
dHand <- data.frame()
##Create hands and remove specified cards from deck, then add them to their respective hands
C1Row <- match(pCard1, OD_Hit$Value)
C1 <- OD_Hit[C1Row,]
pHand <- rbind(pHand, C1)
OD_Hit <- OD_Hit[-C1Row,]

C2Row <- match(pCard2, OD_Hit$Value)
C2 <- OD_Hit[C2Row,]
pHand <- rbind(pHand, C2)
OD_Hit <- OD_Hit[-C2Row,]

C3Row <- match(dCard1, OD_Hit$Value)
C3 <- OD_Hit[C3Row,]
dHand <- rbind(dHand, C3)
OD_Hit <- OD_Hit[-C3Row,]

fullResults <- data.frame()
probList <- data.frame()
numCards <- numDecks*52
for(i in 2:11){
  ##For each possible card value, find the first row that matches that value in the deck dataframe
  ##Add that card to the hand, then create a new deck without the specified card
  rowCard <- match(i, OD_Hit$Value)
  playerHand <- rbind(pHand, OD_Hit[rowCard,])
  drawnCard <- i
  HitOD <- OD_Hit[-rowCard,]
  playerHand <- aceCheck(playerHand)
  pHandT <- sum(playerHand$Value)
  numLeft <- numLeftinDeck(playerHand, dHand, i, numDecks)
  ProbDraw <- numLeft/(numCards-3)
  ##If the player busts after drawing, it is an automatic loss
  if (pHandT>21){
    outcome <- "Loss"
    reason <- "Bust"
    dHandT <- 0
    dRanks <- "Irelevant"
    RowBust <- data.frame(drawnCard, pHandT, outcome, reason, dHandT, dRanks)
    fullResults <- rbind(fullResults, RowBust)
    ProbWinWithCard <- 0
    ProbLoseWithCard <- 1
  } else{
  for(j in 1:2000){
  ##Otherwise, simulate games where the player stands on that hand.
  shufHitOD <- shuffle(HitOD)
  dealerHand <- dHand
  endTurnValues <- dealerTurn(dealerHand, shufHitOD, shoe)
  dealerHand <- endTurnValues[[1]]
  dHandT <- sum(dealerHand$Value)
  dRanks <- compileRanks(dealerHand)
  outcomes <- outcomeHands(pHandT, dHandT)
  outcome <- outcomes[[1]]
  reason <- outcomes[[2]]
  Rowfull <- data.frame(drawnCard, pHandT, outcome, reason, dHandT, dRanks)
  fullResults <- rbind(fullResults, Rowfull)
  }
  WinResults <- fullResults %>% group_by(outcome, drawnCard) %>% summarise(N = n(), 
                                                                         Probability = N/2000) %>% filter(drawnCard == i, outcome=="Win") 
  ProbWinWithCard <- as.double(WinResults[1,4])
  if(pHandT==21){
    ProbLoseWithCard <- 0
  }else{
    LossResults <- fullResults %>% group_by(outcome, drawnCard) %>% summarise(N = n(), 
                                                                         Probability = N/2000) %>% filter(drawnCard == i, outcome=="Loss")
    ProbLoseWithCard <- as.double(LossResults[1,4])
  }
  }
  Expectation <- ProbWinWithCard-ProbLoseWithCard
  PAi <- ProbDraw*Expectation
  Rowi <- data.frame(drawnCard, ProbDraw, Expectation, PAi)
  probList <- rbind(probList, Rowi)
}
colnames(probList) <- c("drawnCard", "ProbDraw", "Expectation", "PAi")
ExpectationHit <- round(sum(probList$PAi), 4)
listRes <- list(ExpectationHit, probList, fullResults)

return(listRes)
}

```

##Fixed Hit 2, one that allows the player to keep hitting based on previously created rows
```{r}
fixedHit2 <- function(pCard1, pCard2, dCard1, numDecks, deckPenetration, curChart){
OD <- makeOD()
OD_Hit <- shuffle(cloneDeck(OD, numDecks))
deck_p <- penetration(OD_Hit, deckPenetration)
OD_Hit <- deck_p[[1]]
shoe <- deck_p[[2]]
pHand <- data.frame()
dHand <- data.frame()
##Create hands and remove specified cards from deck, then add them to their respective hands
C1Row <- match(pCard1, OD_Hit$Value)
C1 <- OD_Hit[C1Row,]
pHand <- rbind(pHand, C1)
OD_Hit <- OD_Hit[-C1Row,]

C2Row <- match(pCard2, OD_Hit$Value)
C2 <- OD_Hit[C2Row,]
pHand <- rbind(pHand, C2)
OD_Hit <- OD_Hit[-C2Row,]

C3Row <- match(dCard1, OD_Hit$Value)
C3 <- OD_Hit[C3Row,]
dHand <- rbind(dHand, C3)
OD_Hit <- OD_Hit[-C3Row,]

fullResults <- data.frame()
probList <- data.frame()
numCards <- numDecks*52
for(i in 2:11){
  ##For each possible card value, find the first row that matches that value in the deck dataframe
  ##Add that card to the hand, then create a new deck without the specified card
  rowCard <- match(i, OD_Hit$Value)
  pHandNew <- rbind(pHand, OD_Hit[rowCard,])
  drawnCard <- i
  HitOD <- OD_Hit[-rowCard,]
  pHandNew <- aceCheck(pHandNew)
  pHandT <- sum(pHandNew$Value)
  numLeft <- numLeftinDeck(pHandNew, dHand, i, numDecks)
  ProbDraw <- numLeft/(numCards-3)
  ##If the player busts after drawing, it is an automatic loss
  if (pHandT>21){
    outcome <- "Loss"
    reason <- "Bust"
    dHandT <- 0
    dRanks <- "Irelevant"
    RowBust <- data.frame(drawnCard, pHandT, outcome, reason, dHandT, dRanks)
    fullResults <- rbind(fullResults, RowBust)
    Expectation <- -1
    PAi <- ProbDraw*Expectation
    Rowi <- data.frame(drawnCard, ProbDraw, Expectation, PAi)
    probList <- rbind(probList, Rowi)
  } else{
  #Many other totals have many combinations of cards that sum to it, so 1000 is adequate
  for(j in 1:1000){
    playerHand <- pHandNew
    pHandT <- sum(playerHand$Value)
    bPlay <- oPlay(playerHand, dHand, curChart)
    shufHitOD <- shuffle(HitOD)
  #Check if drawn card reached a total we know should be hit
  if(bPlay=="H"){
        while( ((bPlay=="H")==TRUE) && pHandT <= 21){
          endTurnValues <- draw(playerHand, shufHitOD, shoe)
          playerHand <- endTurnValues[[1]]
          shufHitOD <- endTurnValues[[2]]
          shoe <- endTurnValues[[3]]
          playerHand <- aceCheck(playerHand)
          pHandT <- sum(playerHand$Value)
          bPlay <- oPlay(playerHand, dHand, curChart)
        }
  }
  #Check for bust
  if(pHandT>21){
      outcome <- "Loss"
      reason <- "Bust"
      dHandT <- sum(dHand$Value)
      dRanks <- compileRanks(dHand)
  #Otherwise, dealer takes a turn
  }else{
  dealerHand <- dHand
  endTurnValues <- dealerTurn(dealerHand, shufHitOD, shoe)
  dealerHand <- endTurnValues[[1]]
  shufHitOD <- endTurnValues[[2]]
  shoe <- endTurnValues[[3]]
  dHandT <- sum(dealerHand$Value)
  dRanks <- compileRanks(dealerHand)
  outcomes <- outcomeHands(pHandT, dHandT)
  outcome <- outcomes[[1]]
  reason <- outcomes[[2]]
  #Compile all data and add row to final dataset
  Rowfull <- data.frame(drawnCard, pHandT, outcome, reason, dHandT, dRanks)
  fullResults <- rbind(fullResults, Rowfull)
  }
  }
  #Calculate win probability
  WinResults <- fullResults %>% filter(drawnCard == i, outcome=="Win") %>% group_by(outcome, drawnCard) %>% summarise(N = n(), Probability = N/1000)
  ProbWinWithCard <- as.double(WinResults[1,4])
  #Calculate loss probability
  if(pHandT==21){
    ProbLoseWithCard <- 0
  }else{
    LossResults <- fullResults %>% filter(drawnCard == i, outcome=="Loss") %>% group_by(outcome, drawnCard) %>% summarise(N = n(),Probability = N/1000)
    ProbLoseWithCard <- as.double(LossResults[1,4])
  }
  #Calculate expectation and add it to probability dataset
  Expectation <- ProbWinWithCard-ProbLoseWithCard
  PAi <- ProbDraw*Expectation
  Rowi <- data.frame(drawnCard, ProbDraw, Expectation, PAi)
  probList <- rbind(probList, Rowi)
  }
colnames(probList) <- c("drawnCard", "ProbDraw", "Expectation", "PAi")
ExpectationHit <- round(sum(probList$PAi), 4)
listRes <- list(ExpectationHit, probList, fullResults)
}
return(listRes)
}
```

##Making the Basic Strategy Chart
```{r}
BSchartRow <- function(FiltStayX, FiltHitX, pTotal){
  fRow <- data.frame()
  fRow[1,1] <- pTotal
  numCombos <- nrow(FiltStayX)
for(i in 2:11){
    weightedProbsStay <- sum(FiltStayX[,i])*(1/numCombos)
    weightedProbsHit <- sum(FiltHitX[,i])*(1/numCombos)
 if(weightedProbsStay  < weightedProbsHit){
   fRow[1,i] <- "H"
 }else if (weightedProbsStay  > weightedProbsHit){
   fRow[1,i] <- "S"
 }
}
colnames(fRow) <- c("","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
return(fRow)
}
```

##Function for finding whether a hand should keep hitting or not based on previous BSChart rows
```{r}
oPlay <- function(playHand, dealHand, currentChart){
  pHandT <- sum(playHand$Value)
  if(dealHand[1,]$Value==1){
    D_upcard <- 11
  } else{
    D_upcard <- dealHand[1,]$Value
  }
  
  if(pHandT <= 11){
    play <- "H"
  }else if(pHandT >= 17){
    play <- "S"
  } else{
    Strat <- currentChart %>% filter(playerTotal==pHandT)
    play <- Strat[1,D_upcard]
    }
  return(play)
}
```

##Basic Strategy Charts, according to sources
```{r}
Total <- c(7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17)
a <- c("H","H","H","H","H","H","S","S","S","S","S")
b <- c("H","H","H","H","H","H","S","S","S","S","S")
c <- c("H","H","H","H","H","S","S","S","S","S","S")
d <- c("H","H","H","H","H","S","S","S","S","S","S")
e <- c("H","H","H","H","H","S","S","S","S","S","S")
f <- c("H","H","H","H","H","H","H","H","H","H","S")
g <- c("H","H","H","H","H","H","H","H","H","H","S")
h <- c("H","H","H","H","H","H","H","H","H","H","S")
i <- c("H","H","H","H","H","H","H","H","H","H","S")
j <- c("H","H","H","H","H","H","H","H","H","H","S")
noStratHardOfficial <- data.frame(Total, a, b, c, d, e, f, g, h, i, j)
colnames(noStratHardOfficial)<-c("HardTotal", 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)
sTotal <- c(13, 14, 15, 16, 17, 18, 19, 20)
a <- c("H","H","H","H","H","S","S","S")
b <- c("H","H","H","H","H","S","S","S")
c <- c("H","H","H","H","H","S","S","S")
d <- c("H","H","H","H","H","S","S","S")
e <- c("H","H","H","H","H","S","S","S")
f <- c("H","H","H","H","H","S","S","S")
g <- c("H","H","H","H","H","S","S","S")
h <- c("H","H","H","H","H","H","S","S")
i <- c("H","H","H","H","H","H","S","S")
j <- c("H","H","H","H","H","H","S","S")
noStratSoftOfficial <- data.frame(sTotal, a, b, c, d, e, f, g, h, i, j)
colnames(noStratSoftOfficial)<-c("SoftTotal", 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)

HardTotal <- c(7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17)
a <- c("H","H","H","D","D","H","S","S","S","S","S")
b <- c("H","H","D","D","D","H","S","S","S","S","S")
c <- c("H","H","D","D","D","S","S","S","S","S","S")
d <- c("H","H","D","D","D","S","S","S","S","S","S")
e <- c("H","H","D","D","D","S","S","S","S","S","S")
f <- c("H","H","H","D","D","H","H","H","H","H","S")
g <- c("H","H","H","D","D","H","H","H","H","H","S")
h <- c("H","H","H","D","D","H","H","H","H","H","S")
i <- c("H","H","H","H","D","H","H","H","H","H","S")
j <- c("H","H","H","H","D","H","H","H","H","H","S")
BSHardOfficial <- data.frame(HardTotal, a, b, c, d, e, f, g, h, i, j)
colnames(BSHardOfficial)<-c("HardTotal", 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)
SoftTotal <- c(13, 14, 15, 16, 17, 18, 19, 20)
a <- c("H","H","H","H","D","S","S","S")
b <- c("H","H","H","H","D","D","S","S")
c <- c("D","D","D","D","D","D","S","S")
d <- c("D","D","D","D","D","D","S","S")
e <- c("D","D","D","D","D","D","D","S")
f <- c("H","H","H","H","H","S","S","S")
g <- c("H","H","H","H","H","S","S","S")
h <- c("H","H","H","H","H","H","S","S")
i <- c("H","H","H","H","H","H","S","S")
j <- c("H","H","H","H","H","S","S","S")
BSSoftOfficial <- data.frame(SoftTotal, a, b, c, d, e, f, g, h, i, j)
colnames(BSSoftOfficial)<-c("SoftTotal", 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)

PairOf_ <- c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11)
a <- c("Y","Y","N","N","Y","Y","Y","Y","N","Y")
b <- c("Y","Y","N","N","Y","Y","Y","Y","N","Y")
c <- c("Y","Y","N","N","Y","Y","Y","Y","N","Y")
d <- c("Y","Y","Y","N","Y","Y","Y","Y","N","Y")
e <- c("Y","Y","Y","N","Y","Y","Y","Y","N","Y")
f <- c("Y","Y","N","N","N","Y","Y","N","N","Y")
g <- c("N","N","N","N","N","N","Y","Y","N","Y")
h <- c("N","N","N","N","N","N","Y","Y","N","Y")
i <- c("N","N","N","N","N","N","Y","N","N","Y")
j <- c("N","N","N","N","N","N","Y","N","N","Y")
BSSplitOfficial <- data.frame(PairOf_, a, b, c, d, e, f, g, h, i, j)
colnames(BSSplitOfficial)<-c("PairOf_", 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)
```
```{r}
write.csv(BSSplitOfficial, "BSSplitOfficial.csv")
write.csv(BSSoftOfficial, "BSSoftOfficial.csv")
write.csv(BSHardOfficial, "BSHardOfficial.csv")
write.csv(noStratHardOfficial, "noStratHardOfficial.csv")
write.csv(noStratSoftOfficial, "noStratSoftOfficial.csv")
```

##Function for reading the optimal play for a given situtation for a given strategy chart
```{r}
optimalPlay <- function(playerHand, dealerHand, soft, dDown){
  if(dealerHand[1,]$Value==1){
    dealerUpCard <- 11
  }else{
    dealerUpCard <- dealerHand[1,]$Value
  }
  pTotal <- sum(playerHand$Value)
  if(pTotal<=8){
    bestPlay <- "H"
  } else if(pTotal>=20){
    bestPlay <- "S"
  } else if(playerHand[1,]$Rank=="A" && playerHand[2,]$Rank=="A"){
    bestPlay <- "H"
  }else if(soft==1){
      if(dDown==1){
        if((any(playerHand==11))==TRUE){
          Strats <- BSSoftOfficial %>% filter(SoftTotal==pTotal)
          bestPlay <- Strats[1, dealerUpCard]
        }else if((any(playerHand==11))==FALSE){
          if(pTotal>=17){
            bestPlay<-"S"
          }else{
            Strats <- BSHardOfficial %>% filter(HardTotal==pTotal)
            bestPlay <- Strats[1, dealerUpCard]  
          }
        }
      }else if(dDown==0){
        if((any(playerHand==11))==TRUE){
          Strats <- noStratSoftOfficial %>% filter(SoftTotal==pTotal)
          bestPlay <- Strats[1, dealerUpCard]
        }else if((any(playerHand==11))==FALSE){
          if(pTotal>=17){
            bestPlay<-"S"
          }else{
            Strats <- noStratHardOfficial %>% filter(HardTotal==pTotal)
            bestPlay <- Strats[1, dealerUpCard]  
          }
        }
      }
  }else if(soft==0){
    if(pTotal>=17){
        bestPlay<-"S"
    }else if(dDown==1){
      Strats <- BSHardOfficial %>% filter(HardTotal==pTotal)
      bestPlay <- Strats[1, dealerUpCard]
    }else if(dDown==0){
      Strats <- noStratHardOfficial %>% filter(HardTotal==pTotal)
      bestPlay <- Strats[1, dealerUpCard]  
    }
  }
  return(bestPlay)
  }


splitCheck <- function(playerHand, dealerHand, split){
  if((playerHand[1,]$Rank == playerHand[2,]$Rank) && split==1){
    Strats <- BSSplitOfficial %>% filter(PairOf_==playerHand[1,]$Value)
    sCheck <- Strats[1, dealerHand[1,]$Value]
    return(sCheck)
  }
  else{
    return("N")
  }
  
}


```

##Function that plays a turn based on what the best play was
```{r}
optimalTurn <- function(playerHand, dealerHand, BSDeck, shoe, bestPlay, soft, dDown){
  pHandT <- sum(playerHand$Value)
  dDownCheck <- FALSE
  if(bestPlay=="D"){
      #Double Down, AKA, Hit once and end round
      dDownCheck <- TRUE
      endTurnValues <- draw(playerHand, BSDeck, shoe)
      playerHand <- endTurnValues[[1]]
      playerHand <- aceCheck(playerHand)
      BSDeck <- endTurnValues[[2]]
      shoe <- endTurnValues[[3]]
  }else if(bestPlay=="H"){
        while( ((bestPlay=="H")==TRUE || (bestPlay=="D")==TRUE) && pHandT <= 21){
          endTurnValues <- draw(playerHand, BSDeck, shoe)
          playerHand <- endTurnValues[[1]]
          BSDeck <- endTurnValues[[2]]
          shoe <- endTurnValues[[3]]
          playerHand <- aceCheck(playerHand)
          pHandT <- sum(playerHand$Value)
          bestPlay <- optimalPlay(playerHand, dealerHand, soft, dDown)
        }
    }
 endTurn <- list(playerHand, BSDeck, shoe, dDownCheck) 
 return(endTurn)
}


splitTurn <- function(playerHand, dealerHand, deck, shoe, soft, dDown){
  
playHand1 <- data.frame()
playHand2 <- data.frame()
dDownCheck1 <- FALSE
dDownCheck2 <- FALSE
playHand1 <- rbind(playHand1, playerHand[1,])
playHand2 <- rbind(playHand2, playerHand[2,])

#Deal new hands
  H1draw <- draw(playHand1, deck, shoe)
  playHand1 <- H1draw[[1]]
  deck <- H1draw[[2]]
  shoe <- H1draw[[3]]
  
  H2draw <- draw(playHand2, deck, shoe)
  playHand2 <- H2draw[[1]]
  deck <- H2draw[[2]]
  shoe <- H2draw[[3]]
  
#Play hand 1 first, to conclusion
  if(sum(playHand1$Value) == 21){
    outcome1 <- "Win"
    reason1 <- "Blackjack"
    endTurn1 <- list(playHand1, dealerHand, outcome1, reason1, deck, shoe, dDownCheck1)
  }else if(sum(playHand1$Value) != 21){
      bestPlay <- optimalPlay(playHand1, dealerHand, soft, dDown)
      endTurn1 <- optimalTurn(playHand1, dealerHand, deck, shoe, bestPlay, soft, dDown)
      playHand1 <- endTurn1[[1]]
      deck <- endTurn1[[2]]
      shoe <- endTurn1[[3]]
      dDownCheck1 <- endTurn1[[4]]
  }
  
#Play second hand next, to conclusion
  if(sum(playHand2$Value) == 21){
    outcome2 <- "Win"
    reason2 <- "Blackjack"
    endTurn2 <- list(playHand2, dealerHand, outcome2, reason2, deck, shoe, dDownCheck2)
  }else if(sum(playHand2$Value) != 21){
    bestPlay <- optimalPlay(playHand2, dealerHand, soft, dDown)
    endTurn2 <- optimalTurn(playHand2, dealerHand, deck, shoe, bestPlay, soft, dDown)
    playHand2 <- endTurn2[[1]]
    deck <- endTurn2[[2]]
    shoe <- endTurn2[[3]]
    dDownCheck2 <- endTurn2[[4]]
  }
  
#After both hands concluded, dealer takes their turn
  endTurnValues <- dealerTurn(dealerHand, deck, shoe)
  dealerHand <- endTurnValues[[1]]
  deck <- endTurnValues[[2]]
  shoe <- endTurnValues[[3]]
  dHandT <- sum(dealerHand$Value)
  
#Find outcome of each hand and make a row for each
  pHandT1 <- sum(playHand1$Value)
  pHandT2 <- sum(playHand2$Value)
  if(pHandT1>21){
    outcome1 <- "Loss"
    reason1 <- "Bust"
  }else if(pHandT1 <= 21){
  outcomes1 <- outcomeHands(pHandT1, dHandT)
  outcome1 <- outcomes1[[1]]
  reason1 <- outcomes1[[2]]
  }
  if(pHandT2>21){
    outcome2 <- "Loss"
    reason2 <- "Bust"
  }else if(pHandT2 <= 21){
  outcomes2 <- outcomeHands(pHandT2, dHandT)
  outcome2 <- outcomes2[[1]]
  reason2 <- outcomes2[[2]]
  }
  
  endRound1 <- list(playHand1, dealerHand, outcome1, reason1, deck, shoe, dDownCheck1)
  endRound2 <- list(playHand2, dealerHand, outcome2, reason2, deck, shoe, dDownCheck2)
  
  endRound <- list(endRound1, endRound2)
  return(endRound)
}
```


##Function to play a game following basic strategy, with soft, ddown, and split being variables set to 1 if the respective ruleset distiction is being allowed for. (i.e. if we are simulating a game where the player follows both a soft and hard total chart, but does not double down or split, it would be 1,0,0)
```{r}
simBSGame <- function(numSims, numDecks, soft, dDown, split){
  BasicStrategySimulations <-data.frame()
  ordered_deck <- makeOD()
  newDeck <- cloneDeck(ordered_deck, numDecks)
  shufDeck <- shuffle(newDeck)
  deck_p <- penetration(shufDeck, 0.20)
  shufDeck <- deck_p[[1]]
  shoe <- deck_p[[2]]
for(i in 1:numSims){
    endRound <- BSGame(shufDeck, shoe, soft, dDown, split)
    #Check if the player split
    if(length(endRound)==2){
      ER1 <- endRound[[1]]
      ER2 <- endRound[[2]]
      FR1 <- compileOutcome(ER1)
      FR2 <- compileOutcome(ER2)
      shufDeck <-ER2[[5]]
      shoe <- ER2[[6]]
      dDownCheck1 <- ER1[[7]]
      dDownCheck2 <- ER2[[7]]
      if(dDownCheck1 == TRUE){
      ##If the player doubled down, we double the game, as in real life you double your bet.
      ##so you either win twice as much as you bet or lose double. Meaning whatever the
      ##outcome of the round, if we double it it will impact the probabilities accordingly?
      BasicStrategySimulations <- rbind(BasicStrategySimulations, FR1)
      BasicStrategySimulations <- rbind(BasicStrategySimulations, FR1)
      }else if(dDownCheck1==FALSE){
     BasicStrategySimulations <- rbind(BasicStrategySimulations, FR1)
      }
      if(dDownCheck2 == TRUE){
      BasicStrategySimulations <- rbind(BasicStrategySimulations, FR2)
      BasicStrategySimulations <- rbind(BasicStrategySimulations, FR2)
      }else if(dDownCheck2 ==FALSE){
     BasicStrategySimulations <- rbind(BasicStrategySimulations, FR2)
    }
  }else{
    finalRow <- compileOutcome(endRound)
    shufDeck <-endRound[[5]]
    shoe <- endRound[[6]]
    dDownCheck <- endRound[[7]]
    if(dDownCheck == TRUE){
      BasicStrategySimulations <- rbind(BasicStrategySimulations, finalRow)
      BasicStrategySimulations <- rbind(BasicStrategySimulations, finalRow)
      }else{
     BasicStrategySimulations <- rbind(BasicStrategySimulations, finalRow)
      }
  }
  if(nrow(shufDeck)==0){
    shufDeck <- shuffle(newDeck)
    deck_p <- penetration(shufDeck, 0.20)
    shufDeck <- deck_p[[1]]
    shoe <- deck_p[[2]]
  }
}
return(BasicStrategySimulations)
}

BSGame <- function(deck, shoe, soft, dDown, split){
  #Make new hands and set checks to default
  dDownCheck <- FALSE
  dDownCheck1 <- FALSE
  dDownCheck2  <- FALSE
  playerHand <- data.frame()
  dealerHand <- data.frame()
  playHand1 <- data.frame()
  playHand2 <- data.frame()
  ## Deal
  dealtHands <- deal(playerHand, dealerHand, deck, shoe)
  playerHand <- dealtHands[[1]]
  dealerHand <- dealtHands[[2]]
  BSDeck <- dealtHands[[3]]
  shoe <- dealtHands[[4]]
  pHandT <- sum(playerHand$Value)
  dHandT <- sum(dealerHand$Value)
  ## Blackjack check
  if(pHandT==21){
    if(dHandT==21){
      outcome <- "Push"
      reason <- "Double Blackjacks"
      endRound <- list(playerHand, dealerHand, outcome, reason, BSDeck, shoe, dDownCheck)
    }else{
      outcome <- "Win"
      reason <- "Blackjack"
      endRound <- list(playerHand, dealerHand, outcome, reason, BSDeck, shoe, dDownCheck)
    }
  }else if(dHandT==21){
    outcome <- "Loss"
    reason <- "Dealer Blackjack"
    endRound <- list(playerHand, dealerHand, outcome, reason, BSDeck, shoe, dDownCheck)
##Player turn, based on allowed options and strategy chart
  }else{
    sCheck <- splitCheck(playerHand, dealerHand, split)
    if(sCheck=="Y"){
      endRound <- splitTurn(playerHand, dealerHand, BSDeck, shoe, soft, dDown)
    }
    #if no split, play hand normally
    else{
      if(pHandT<=8){
        bestPlay <- "H"
      }else if(pHandT>=20){
        bestPlay <- "S"
      }else {
        bestPlay <- optimalPlay(playerHand, dealerHand, soft, dDown)
      }
    endTurn <- optimalTurn(playerHand, dealerHand, BSDeck, shoe, bestPlay, soft, dDown)
    playerHand <- endTurn[[1]]
    BSDeck <- endTurn[[2]]
    shoe <- endTurn[[3]]
    dDownCheck <- endTurn[[4]]
    pHandT <- sum(playerHand$Value)
    #Bust check
    if(pHandT>21){
      outcome <- "Loss"
      reason <- "Bust"
    }else{
    #Dealer turn
    endTurnValues <- dealerTurn(dealerHand, BSDeck, shoe)
    dealerHand <- endTurnValues[[1]]
    BSDeck <- endTurnValues[[2]]
    shoe <- endTurnValues[[3]]
    dHandT <- sum(dealerHand$Value)
    #Outcomes
    outcomes <- outcomeHands(pHandT, dHandT)
    outcome <- outcomes[[1]]
    reason <- outcomes[[2]]
    }
    endRound <- list(playerHand, dealerHand, outcome, reason, BSDeck, shoe, dDownCheck)
    }
  }
  return(endRound)
  }

```


##Misc Functions
```{r}
compileOutcome <- function(outcomeList){
  
  playerHand <- outcomeList[[1]]
  pRanks <- compileRanks(playerHand)
  pTotal <- sum(playerHand$Value)
  
  dealerHand <- outcomeList[[2]]
  dRanks <- compileRanks(dealerHand)
  dTotal <- sum(dealerHand$Value)
  
  outcome <- outcomeList[[3]]
  reason <-outcomeList[[4]]
  if(length(outcomeList)>4){
    deck <- outcomeList[[5]]
    shoe <- outcomeList[[6]]
    remainDeck <- nrow(deck)
    ShoeSize <- nrow(shoe)
    finalData <- data.frame(outcome, reason, pTotal, dTotal, pRanks, dRanks, remainDeck, ShoeSize)
  }else{
    finalData <- data.frame(outcome, reason, pTotal, dTotal, pRanks, dRanks)
  }
  return(finalData)
}

compileRanks <- function(hand){
    numCards <- nrow(hand)
  if(numCards==1){
    Ranks <- hand[1,1]
  }else{
  for(i in 1:numCards){
    if(i==1){
      Ranks <- hand[1,1]
    } else {
    ranki <- hand[i, 1]
    Ranks <- paste(Ranks, ranki, sep = ", ")
  }
  }
  }
  return(Ranks)
}

removeValue <- function(value, ordered_deck){
  rowCard <- match(value, ordered_deck$Value)
  ordered_deck <- ordered_deck[-rowCard,]
}

numLeftinDeck <- function(phand, dhand, cardvalue, numDecks){
  ##There 16 cards valued at 10, and 4 of everything else, in a single deck
if(cardvalue==10){
  numLeft <- 16*numDecks
}else{
  numLeft <- 4*numDecks
}
if(phand[1,]$Value==cardvalue){
  numLeft <- numLeft-1
}
if(phand[2,]$Value==cardvalue){
  numLeft <- numLeft-1
}
if(dhand[1,]$Value==cardvalue){
  numLeft <- numLeft-1
}
  return(numLeft)
}
```
