---
title: "Simulations"
author: "Bodhi Bourland"
date: "2024-02-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(knitr)
library(gridExtra)
```

## Mimic the dealer

```{r, include=FALSE}
mimicStats <- simMimicDealer(100000, 6)
MoutcomeResults <- mimicStats %>% group_by(outcome) %>% summarise(n=n(),                                                        Probability = n/nrow(mimicStats))
row6Deck<- c("6", MoutcomeResults[1,3], MoutcomeResults[2,3], MoutcomeResults[3,3])
```
```{r}
write.csv(mimicStats, "MimicTheDealerSim.csv", row.names=FALSE)
```


## Proving you should always Stand on 17

```{r, include=FALSE}
##Probabilities after staying
Stay17 <- data.frame()
for(i in 2:11){
sim1 <- fixedStay(7, 10, i, 6, 0.20)
sim2 <- fixedStay(8, 9, i, 6, 0.20)
Stay17[1,i-1] <- sim1[[1]]*100
Stay17[2,i-1] <- sim2[[1]]*100
}
colnames(Stay17) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("7, 10", "8, 9")
Stay17 <- data.frame(Combination, Stay17)
```

```{r, include=FALSE}
##Probabilities after hitting
Hit17 <- data.frame()
for(i in 2:11){
  sim1 <- fixedHit(7, 10, i, 6, 0.20)
  sim2 <- fixedHit(8, 9, i, 6, 0.20)
  Hit17[1,i-1] <- sim1[[1]]*100
  Hit17[2,i-1] <- sim2[[1]]*100
}
colnames(Hit17) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("7, 10", "8, 9")
Hit17 <- data.frame(Combination, Hit17)
```

```{r}
bsBottomRow <- BSchartRow(Stay17, Hit17, 17)
colnames(bsBottomRow) <- c("playerTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
write.csv(Stay17, "Stay17.csv")
write.csv(Hit17, "Hit17.csv")
```

##Starting Boundaries of Basic Strategy Chart
```{r}
begChart <- data.frame()
for(i in 1:5){
  begChart[i,1] <- i+6
  for(j in 2:11){
    begChart[i,j] <- "H"
  }
colnames(begChart) <- c("playerTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
beginningChart <- rbind(begChart, bsBottomRow)

}
begChart$playerTotal <- as.integer(begChart$playerTotal)
bsBottomRow$playerTotal <- as.integer(bsBottomRow$playerTotal)
write.csv(beginningChart, "boundariesBSChart.csv")
```

## Basic Strategy: Player total = 16

```{r, include=FALSE}
Stay16 <- data.frame()
for(i in 2:11){
sim1 <- fixedStay(6, 10, i, 6, 0.2)
sim2 <- fixedStay(7, 9, i, 6, 0.2)
Stay16[1,i-1] <- sim1[[1]]*100
Stay16[2,i-1] <- sim2[[1]]*100
}
colnames(Stay16) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("6, 10", "7, 9")
Stay16 <- data.frame(Combination, Stay16)
```


```{r, include=FALSE}
Hit16 <- data.frame()
for(i in 2:11){
  sim1 <- fixedHit(6, 10, i, 6, 0.2)
  sim2 <- fixedHit(7, 9, i, 6, 0.2)
  Hit16[1,i-1] <- sim1[[1]]
  Hit16[2,i-1] <- sim2[[1]]
}
colnames(Hit16) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("6, 10", "7, 9")
Hit16 <- data.frame(Combination, Hit16)
```


```{r}
BSrow16<- BSchartRow(Stay16, Hit16, 16)
colnames(BSrow16) <- c("playerTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
write.csv(Stay16, "Stay16.csv")
write.csv(Hit16, "Hit16.csv")
write.csv(BSrow16, "Row16.csv")
curChart <- rbind(begChart, BSrow16)
BSrow16 <- BSrow16[,-1]
BSrow16
```

##PTotal == 15
```{r}
Stay15 <- data.frame()
for(i in 2:11){
sim1 <- fixedStay(5, 10, i, 6, 0.2)
sim2 <- fixedStay(6, 9, i, 6, 0.2)
sim3 <- fixedStay(7, 8, i, 6, 0.2)
Stay15[1,i-1] <- sim1[[1]]
Stay15[2,i-1] <- sim2[[1]]
Stay15[3,i-1] <- sim3[[1]]
}
colnames(Stay15) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("5, 10", "6, 9", "7, 8")
Stay15 <- data.frame(Combination, Stay15)
```
```{r}
Hit15 <- data.frame()

for(i in 2:11){
  print("sim1")
  sim1 <- fixedHit2(5, 10, i, 6, 0.2, curChart)
  print("sim2")
  sim2 <- fixedHit2(6, 9, i, 6, 0.2, curChart)
  print("sim3")
  sim3 <- fixedHit2(7, 8, i, 6, 0.2, curChart)
  Hit15[1,i-1] <- sim1[[1]]
  Hit15[2,i-1] <- sim2[[1]]
  Hit15[3,i-1] <- sim3[[1]]
}
colnames(Hit15) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("5, 10", "6, 9", "7, 8")
Hit15 <- data.frame(Combination, Hit15)
colnames(Hit15) <- c("Combination","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Hit15
```
```{r}
BSrow15<- BSchartRow(Stay15, Hit15, 15)
colnames(BSrow15) <- c("playerTotal" ,"2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
BSrow15
Hit15 <- Hit15[,-1]
```
```{r}
write.csv(BSrow15, "Row15.csv")
curChart <- rbind(begChart, BSrow15, BSrow16)
```

##PTotal == 14
```{r}
Stay14 <- data.frame()
for(i in 2:11){
sim1 <- fixedStay(4, 10, i, 6, 0.2)
sim2 <- fixedStay(5, 9, i, 6, 0.2)
sim3 <- fixedStay(6, 8, i, 6, 0.2)
Stay14[1,i-1] <- sim1[[1]]
Stay14[2,i-1] <- sim2[[1]]
Stay14[3,i-1] <- sim3[[1]]
}
colnames(Stay14) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("4, 10", "5, 9", "6, 8")
Stay14 <- data.frame(Combination, Stay14)
colnames(Stay14) <- c("Combination","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
```
```{r}
Hit14 <- data.frame()
for(i in 2:11){
  sim1 <- fixedHit2(4, 10, i, 6, 0.2, curChart)
  sim2 <- fixedHit2(5, 9, i, 6, 0.2, curChart)
  sim3 <- fixedHit2(6, 8, i, 6, 0.2, curChart)
  Hit14[1,i-1] <- sim1[[1]]
  Hit14[2,i-1] <- sim2[[1]]
  Hit14[3,i-1] <- sim3[[1]]
}
colnames(Hit14) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("4, 10", "5, 9", "6, 8")
Hit14 <- data.frame(Combination, Hit14)
colnames(Hit14) <- c("Combination","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
```
```{r}
BSrow14<- BSchartRow(Stay14, Hit14, 14)
colnames(BSrow14) <- c("playerTotal" ,"2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
write.csv(BSrow14, "Row14.csv")
curChart <- rbind(begChart, BSrow14, BSrow15, BSrow16)
```

##PTotal == 13
```{r}
Stay13 <- data.frame()
for(i in 2:11){
sim1 <- fixedStay(3, 10, i, 6, 0.2)
sim2 <- fixedStay(4, 9, i, 6, 0.2)
sim3 <- fixedStay(5, 8, i, 6, 0.2)
sim4 <- fixedStay(6, 7, i, 6, 0.2)
Stay13[1,i-1] <- sim1[[1]]
Stay13[2,i-1] <- sim2[[1]]
Stay13[3,i-1] <- sim3[[1]]
Stay13[4,i-1] <- sim4[[1]]
}
colnames(Stay13) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("3, 10", "4, 9", "5, 8", "6, 7")
Stay13 <- data.frame(Combination, Stay13)
```
```{r}
Hit13 <- data.frame()
for(i in 2:11){
  sim1 <- fixedHit2(3, 10, i, 6, 0.2, curChart)
  sim2 <- fixedHit2(4, 9, i, 6, 0.2, curChart)
  sim3 <- fixedHit2(5, 8, i, 6, 0.2, curChart)
  sim4 <- fixedHit2(6, 7, i, 6, 0.2, curChart)
  Hit13[1,i-1] <- sim1[[1]]
  Hit13[2,i-1] <- sim2[[1]]
  Hit13[3,i-1] <- sim3[[1]]
  Hit13[4,i-1] <- sim4[[1]]
}
colnames(Hit13) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("3, 10", "4, 9", "5, 8", "6, 7")
Hit13 <- data.frame(Combination, Hit13)
colnames(Hit13) <- c("Combination","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
```
```{r}
BSrow13<- BSchartRow(Stay13, Hit13, 13)
colnames(BSrow13) <- c("playerTotal" ,"2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
write.csv(BSrow13, "Row13.csv")
curChart <- rbind(begChart, BSrow13, BSrow14, BSrow15, BSrow16)
```


##PTotal == 12
```{r}
Stay12 <- data.frame()
for(i in 2:11){
sim1 <- fixedStay(2, 10, i, 6, 0.2)
sim2 <- fixedStay(3, 9, i, 6, 0.2)
sim3 <- fixedStay(4, 8, i, 6, 0.2)
sim4 <- fixedStay(5, 7, i, 6, 0.2)
Stay12[1,i-1] <- sim1[[1]]
Stay12[2,i-1] <- sim2[[1]]
Stay12[3,i-1] <- sim3[[1]]
Stay12[4,i-1] <- sim4[[1]]
}
colnames(Stay12) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("2, 10", "3, 9", "4, 8", "5, 7")
Stay12 <- data.frame(Combination, Stay12)
colnames(Stay12) <- c("Combination","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
```


##PTotal=12
```{r}
Hit12 <- data.frame()
for(i in 2:11){
  sim1 <- fixedHit2(2, 10, i, 6, 0.2, curChart)
  sim2 <- fixedHit2(3, 9, i, 6, 0.2, curChart)
  sim3 <- fixedHit2(4, 8, i, 6, 0.2, curChart)
  sim4 <- fixedHit2(5, 7, i, 6, 0.2, curChart)
  Hit12[1,i-1] <- sim1[[1]]
  Hit12[2,i-1] <- sim2[[1]]
  Hit12[3,i-1] <- sim3[[1]]
  Hit12[4,i-1] <- sim4[[1]]
}
colnames(Hit12) <- c("2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
Combination <- c("2, 10", "3, 9", "4, 8", "5, 7")
Hit12 <- data.frame(Combination, Hit12)
colnames(Hit12) <- c("Combination","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
```

```{r}
BSrow12<- BSchartRow(Stay12, Hit12, 12)
colnames(BSrow12) <- c("playerTotal" ,"2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
write.csv(BSrow12, "Row12.csv")
```

##Simulating Soft Hands
```{r}
SoftChart <- data.frame("SoftTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
colnames(SoftChart) <- c("SoftTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
for(j in 1:8){
SoftChart[i, 1] <- i+12
StayX <- data.frame("SoftTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
colnames(StayX) <- c("SoftTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
HitX <- data.frame("SoftTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
colnames(HitX) <- c("SoftTotal","2", "3", "4", "5", "6", "7", "8", "9", "10", "A")
StayX[j, 1] <- j+12
HitX[j, 1] <- j+12
for(i in 2:11){
simStay <- fixedStay(11, j+1, i)
StayX[j,i] <- simStay[[1]]
simHit <- fixedHit(11, j+1, i)
HitX[j,i] <- simHit[[1]]
}
BSrowX<- BSchartRow(StayX, HitX, j+12)
SoftChart <- rbind(SoftChart, BSrowX) 
}
```

##Simulating Basic Strategy, only hit and Stand, disregard if hand is hard or soft
```{r}
onlyHitStandNoSoft <- simBSGame(100000, 6, soft = 0, dDown = 0, split = 0)
Sim_HitStand <- onlyHitStandNoSoft %>% group_by(outcome) %>% summarise(numGames=n(),                                                        Probability = n/nrow(onlyHitStandNoSoft))
Expectation <- Sim_HitStand[3,3]-Sim_HitStand[1,3]
Expectation
```
```{r}
-0.0553
write.csv(onlyHitStandNoSoft, "BS_Control.csv", row.names=FALSE)
```

##Simulating Basic Strategy, only hit and Stand, difference between soft and hard totals.

```{r}
onlyHitStandwithSoft <- simBSGame(100000, 6, soft = 1, dDown = 0, split = 0)
Sim_HitStandSoft <- onlyHitStandwithSoft %>% group_by(outcome) %>% summarise(numGames=n(),                                                        Probability = n/nrow(onlyHitStandwithSoft))
Expectation <- Sim_HitStandSoft[3,3]-Sim_HitStandSoft[1,3]
Expectation
```
```{r}
-0.0528
0.0025 gained
write.csv(onlyHitStandwithSoft, "BS_SoftIncluded.csv", row.names=FALSE)
```

##Simulating Basic Strategy, allowing all options/distictions expect splitting

```{r}
dDownwithSoft <- simBSGame(100000, 6, soft = 1, dDown = 1, split = 0)
Sim_dDownandSoft <- dDownwithSoft %>% group_by(outcome) %>% summarise(numGames=n(),                                                        Probability = n/nrow(dDownwithSoft))
```
```{r}
-0.0586
write.csv(dDownwithSoft, "BS_dDown_SoftIncluded.csv", row.names=FALSE)
```

```{r}
dDownNoSoft <- simBSGame(100000, 6, soft = 0, dDown = 1, split = 0)
Sim_dDownOnly <- dDownNoSoft %>% group_by(outcome) %>% summarise(numGames=n(),                                                        Probability = n/nrow(dDownNoSoft))
Expectation <- Sim_dDownOnly[3,3]-Sim_dDownOnly[1,3]
Expectation
```
```{r}
-0.0448
0.0105 gained
write.csv(dDownNoSoft, "BS_dDown_NoSoft.csv", row.names=FALSE)
```

#Simulating Basic Strategy, Adding splits but no soft distinction or double downs
```{r}
onlyHitStandSplit <- simBSGame(100000, 6, soft = 0, dDown = 0, split = 1)
Sim_HitStandSplit <- onlyHitStandSplit %>% group_by(outcome) %>% summarise(numGames=n(), Probability = n/nrow(onlyHitStandSplit))
Expectation <- Sim_HitStandSplit[3,3]-Sim_HitStandSplit[1,3]
Expectation
```
```{r}
-0.0548
0.0005 gained
write.csv(onlyHitStandSplit, "BS_SplitOnly.csv", row.names=FALSE)
```

##Simulating Basic Strategy, allowing all distictions

```{r}
FullBS <- simBSGame(100000, 6, soft = 1, dDown = 1, split = 1)
Sim_FullBasicStrategy <- FullBS %>% group_by(outcome) %>% summarise(numGames=n(),                                                        Probability = numGames/nrow(FullBS))
Expectation <- Sim_FullBasicStrategy[3,3]-Sim_FullBasicStrategy[1,3]
Expectation
```

```{r}
-0.0239
6.04 gained from mimick
write.csv(FullBS, "FullBS.csv", row.names=FALSE)
```

```{r}

Sim_HitStand <- BS_Control %>% group_by(outcome) %>% summarise(numGames =n(),Probability = round(numGames/nrow(BS_Control)*100,2))

Sim_HitStandSoft <- onlyHitStandwithSoft %>% group_by(outcome) %>% summarise(numGames=n(),Probability = round(numGames/nrow(onlyHitStandwithSoft)*100,2))

Sim_dDownOnly <- dDownNoSoft %>% group_by(outcome) %>% summarise(numGames=n(),Probability = round(numGames/nrow(dDownNoSoft)*100,2))

Sim_SplitOnly <- onlyHitStandSplit %>% group_by(outcome) %>% summarise(numGames=n(), Probability = round(numGames/nrow(onlyHitStandSplit)*100,2))

Sim_FullBasicStrategy <- FullBS %>% group_by(outcome) %>% summarise(numGames=n(), Probability = round(numGames/nrow(FullBS)*100,2))

write.csv(Sim_HitStand, "Sim_HitStand.csv", row.names=FALSE)
write.csv(Sim_HitStandSoft, "Sim_HitStandSoft.csv", row.names=FALSE)
write.csv(Sim_dDownOnly, "Sim_dDownOnly.csv", row.names=FALSE)
write.csv(Sim_SplitOnly, "Sim_SplitOnly.csv", row.names=FALSE)
write.csv(Sim_FullBasicStrategy, "Sim_FullBasicStrategy.csv", row.names=FALSE)
```
